### 2017-10-27 src/4_OnlineLearning.tex
Chapter 3 reviewed

Regularization can be understood using two contexts: learning theory
(probabilistic) and inverse problems (deterministic). In the context of
learning, regularization refers to techniques allowing to avoid over-fitting and
the desire property of the selected estimator is to perform well on new data (to
generalise), e.g. regularized least squares. In the context of inverse problems,
regularization objective is to stabilise, with respect to noise, a possibly
ill-conditioned matrix inversion problem e.g spectral cut-off and Tikhonov
regularization or ridge regression (RR) \cite{ tikhonov1977}.

@@

\begin{equation*}
\label{eq:optsolRR}
\mathbf{\mathbf{X}}_*=(\mathbf{A}^\top \mathbf{A}+\lambda \mathbb{I})^{-1}\mathbf{A}^\top y \, ,
\end{equation*}

@@

The major difference between SVM and ANN is in the error optimisation. In ANN, the aim of learning is to obtain a set of weight values which minimise the training error 
minimum while training adjust the capacity of the machine.

@@

In batch learning, the training phase is commonly a computationally expensive
process. Therefore, when new data arrives it can't be included easily to the
model. Moreover, it could happen that we won't have enough time to process the
new data before more data arrives. In online learning algorithms there is no
training phase, but the model is updated and evaluated at every time step. This
model updating is computationally less expensive than a training phase.

@@

The objective is minimize this loss function for all instances.
The quality of online learning algorithms is measured by a quantity known as
regret which is the difference between the performance of the online algorithm
and its optimal predictor $E^* \in \Theta$ given by:

@@

If we define $\displaystyle\mathbf{S}= \sum_{t=1}^{m+1} \mathbf{a}_t
\mathbf{a}_t  ^\intercal + \gamma \mathbb{I} $ and $\mathbf{W}=
\displaystyle\sum_{t=1}^m \mathbf{a}_t \mathbf{y}_t$, the
algorithm~\ref{alg:AAR} is slightly different to the algorithm~\ref{alg:RR}, 
which updated matrix $\mathbf{S}$ before making the prediction:

@@

\subsection{Efficient computation}
In regression problems we always require getting a matrix inverse which is
computationally expensive. However, in stream data problems there is a way to
obtain a matrix inverse approximation using the Sherman-Morrison-Woodbury. This
formula allows to get the inverse of a matrix $\mathbf{A+uv^\top}$ if we previously calculated the
inverse of $\mathbf{A}$ as follows:

\begin{equation}
\label{eq:SMW}
(\mathbf{A+uv^\top})^{-1}=\mathbf{A}^{-1}-
\frac{\mathbf{A^{-1}uv^\top A^{-1}}}{\mathbf{1+v^\top A^{-1}u}}
\end{equation}

Alternatively, Coleman and
Sun~\cite{coleman+sun2010} presented an iterative algorithm which
uses $\mathbf{X}(\lambda)$ to approximate $\mathbf{X}(0)$.

Using the compact SVD (shown in equation~(\ref{eq:compactsvd}))
$\mathbf{X}(\lambda)$ is expressed as follows:

\begin{eqnarray}
\label{eq:optsolRRsvd}
\mathbf{X}(\lambda) & = & (\mathbf{A}^\top \mathbf{A}+ \lambda
\mathbb{I})^{-1}\mathbf{A}^\top \mathbf{Y} \nonumber \\
& = &\mathbf{V}_1(\Sigma_1^2+\lambda \mathbb{I})^{-1}\mathbf{\Sigma_1
U_1^\top Y}
\end{eqnarray}

\noindent where is easy to see that $\mathbf{X}(\lambda) \rightarrow
\mathbf{\hat{X}}=\mathbf{V}_1 \mathbf{\Sigma}_1^{-1}\mathbf{U}_1^\top
\mathbf{Y}$ as $\lambda \rightarrow 0$.

The method consists in obtaining $\mathbf{X}(\lambda)$ and then refine
by adding more terms of its Taylor expansion to approximate
$\mathbf{X}(0) = \mathbf{\hat{X}}$. The Taylor expansion about $\lambda_0$ is:

\begin{equation}
\label{eq:taylor}
    \mathbf{X}(\lambda)=\mathbf{X}(\lambda_0) + \sum_{k=1}^\infty
    \mathbf{s}_k(\lambda-\lambda_0)^{k}
\end{equation}

\noindent where $\mathbf{s}_k=\frac{1}{k!}\mathbf{X}(\lambda)^{(k)}$
and $\mathbf{X}(\lambda)^{(k)}$ is obtained by taking differences 
$\frac{\partial}{\partial \lambda}$ of:

\begin{eqnarray*}
(\mathbf{A}^\top \mathbf{A}+ \lambda\mathbb{I}) \mathbf{X}(\lambda) & = & \mathbf{A}^\top \mathbf{Y}\\
(\mathbf{A}^\top \mathbf{A}+ \lambda\mathbb{I}) \mathbf{X}(\lambda)^{(1)} + \mathbf{X}(\lambda)& = & 0 \\
\mathbf{X}(\lambda)^{(1)}  &=& -(\mathbf{A}^\top \mathbf{A}+ \lambda\mathbb{I}) ^{-1} \mathbf{X}(\lambda) \\
\mathbf{X}(\lambda)^{(2)}  &=& -2(\mathbf{A}^\top \mathbf{A}+ \lambda\mathbb{I}) ^{-1} \mathbf{X}(\lambda)^{(1)} \\
& \vdots & \\
\mathbf{X}(\lambda)^{(k)}  &=& -k(\mathbf{A}^\top \mathbf{A}+ \lambda\mathbb{I}) ^{-1} \mathbf{X}(\lambda)^{(k-1)} 
\end{eqnarray*}

\noindent for $\lambda=0$ we have:

\begin{equation}
\label{eq:taylor}
    \mathbf{X}(0)=\mathbf{X}(\lambda_0) + \sum_{k=1}^\infty
     (-1)^k \mathbf{s}_k \lambda_0^k
\end{equation}

\noindent therefore in order to ensure convergence, we can see that$\lambda_0$
selection cannot be large.

The algorithm for computing $\mathbf{X}(0)$ is the following:

\begin{algorithm}[H]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\mathbf{A}$: design matrix \\
$\mathbf{Y}$: response matrix \\
$\lambda$: rank deficient parameter \\
\ENSURE  $\,$ \\
$\mathbf{X}$: parameters \\
\STATE $\mathbf{M}=\mathbf{A^\top A}$ \\
\STATE Initialize $\mathbf{Q R}=\mathbf{M}$ \\
\STATE $\mathbf{X} = \mathbf{R^{-1}Q^\top Y}$ \\
\STATE $\mathbf{s} = \mathbf{X}$ \\
\FOR { $i = 1,2,3,\dots$ }
	\STATE $\mathbf{s} =
        -(\mathbf{M}+\lambda\mathbb{I})^{-1}\mathbf{s}$\\
	\STATE $\mathbf{X}=\mathbf{X} + (-1)^i {\color{red}\mathbf{s}
        \lambda^i}$
\ENDFOR
\end{algorithmic}
\caption{Algorithm for handling rank deficient matrices}
\label{alg:coleman}
\end{algorithm}

The algorithm~\ref{alg:coleman} solves equation~(\ref{eq:taylor}). However, this
version is unstable since typically $\|\mathbf{s}\|$ is very large and
$\lambda^i$ is very small ($\lambda$ is small).

The following algorithm shows a more stable version of
algorithm~\ref{alg:coleman}.

\begin{algorithm}[H]
\begin{algorithmic}[1]
\REQUIRE $\,$ \\
$\mathbf{A}$: design matrix \\
$\mathbf{Y}$: response matrix \\
$\lambda$: rank deficient parameter \\
\ENSURE  $\,$ \\
$\mathbf{X}$: parameters \\
\STATE $\mathbf{M}=\mathbf{A^\top A}$ \\
\STATE Initialize $\mathbf{Q R}=\mathbf{M}$ \\
\STATE $\mathbf{X} = \mathbf{R^{-1}Q^\top Y}$ \\
\STATE $\mathbf{t} = \mathbf{X}$ \\
\FOR { $i = 1,2,3,\dots$ }
        \STATE $\mathbf{t} =\lambda \mathbf{t}$  
        \STATE $\mathbf{t} =  -(\mathbf{M}+\lambda\mathbb{I})^{-1}\mathbf{t}$
	\STATE $\mathbf{X}=\mathbf{X} + \mathbf{t}$
\ENDFOR
\end{algorithmic}
\caption{Algorithm for handling rank deficient matrices improved}
\label{alg:colemanimproved}
\end{algorithm}

Both algorithms are equivalent, but algorithm~\ref{alg:colemanimproved} is more
stable and converges in typically less than 10 steps.
It is important to notice that the QR factorisation is computed only once and it
is computationally less expensive than the SVD.

%Despite the bias-variance tradeoff provides a conceptual framework for
%determining a good model, is not directly useful. Some popular methods
%for determine model selection are:
%
%\begin{itemize}
%    \item Akaike information criterion (AIC) (Akaike, 1974)
%    \item Bootstrap-based selection (Efron and Tibshirani, 1997)
%    \item Cross-validation (Stone, 1974)
%\end{itemize}
%
%On the other hand, $\lambda$ can also be used to reduce computational time as it
%is shown in equation~(\ref{eq:taylor}). Golub (~\cite{golub1965}) suggested that
%$\lambda$ be chosen so that:
%
%\[
%\frac{\lambda}{\lambda + \delta^2} < 0.1
%\]
%
%\noindent where $\delta$ is a lower bound of the smallest non-zero singular
%value $\sigma_k$. One approach is to choose the greatest lower bound for
%$\delta=\sigma_k$. Hence:
%
%\[
%\frac{\lambda}{\lambda + \sigma_k^2} < 0.1
%\]
%
%\noindent if we choose $\lambda=\beta \sigma_k^2$ then we get the following
%relation:
%
%
%\[
%\frac{\beta \sigma_k^2}{\beta \sigma_k^2 + \sigma_k^2} =
%\frac{\beta}{\beta+1} < 0.1 \Rightarrow \beta < \frac{1}{9}  
%\]
%
%Experiments done by ~\cite{coleman+sun2010} have shown that $\beta = 0.01$
%produces satisfactory results. This means that $\lambda = 0.01 \sigma_k^2$.
%
%However, get $\sigma_k$ imply obtaining first the SVD, which is
%computational expensive.
%
%Since the algorithm presented by ~\cite{coleman+sun2010} already compute the QR
%factorization of the matrix $\mathbf{A}$ they show a procedure for getting
%a $\lambda$ approximation using QR.
%
%\begin{enumerate}
%\item Compute QR factorization: $\mathbf{A} = \mathbf{Q_1R_1}$
%\item Let $\mathbf{W}$ denote the set of absolute values of the nonzero diagonal
%elements of $\mathbf{R_1}$. Let $w_{\text{min}}$ and $w_{\text{max}}$ denote the
%smallest and largest elements of $\mathbf{W}$ respectively. Both 
%$\lambda_1 = \hat{\beta} w_{\text{min}}^2$
%$\lambda_2 = \hat{\beta} \frac{w_{\text{min}}^2}{w_{\text{max}}^2}$ where
%$\hat{\beta}=0.00025$ produce satisfactory results.
%\item The $\lambda_{\text{QR}}$ approximation is obtained as:
%$\lambda_{\text{QR}} = \frac{\lambda_1 + \lambda_2}{2}$  
%\end{enumerate}

@@

Forecast performance is evaluated using different methods. We have chosen three
measures usually used:
\begin{description}
\item[MAPE] Mean Average Percent Error which presents forecast errors as a
percentage.
\begin{equation}\label{eq:MAPE}
\text{MAPE} = \frac{1}{N} \sum_{t=1}^{N} 
\frac{\left|\mathbf{y}_t-\hat{\mathbf{y}}_t\right|}{\left|\mathbf{y}_t\right|}
 \times 100 
\end{equation}
\item[MAE] Mean Average Error which measures the distance between forecasts to the
true value.
\begin{equation}\label{eq:MAE}
\text{MAE} = \frac{1}{N} \sum_{t=1}^{N} 
\left| 
\mathbf{y}_t-\hat{\mathbf{y}}_t
\right| 
\end{equation}
\item[MSE]  Mean Square Error measures the distance between forecasts
and the true values and large deviations from the true value have a
large impact due to squaring forecast error.
\begin{equation}\label{eq:MSE}
\text{MSE} = 
\frac{\displaystyle \sum_{t=1}^{N} (\mathbf{y}_t-\hat{\mathbf{y}}_t)^2}{N}
\end{equation}
\item[RMSE] Root Mean Square Error also measures the distance between forecasts
to the true values but, unlike MAE, large deviations from the true value have a
large impact on RMSE due to squaring forecast error.
\begin{equation}\label{eq:RMSE}
\text{RMSE} = \sqrt{
\frac{\displaystyle \sum_{t=1}^{N} (\mathbf{y}_t-\hat{\mathbf{y}}_t)^2}{N}}
\end{equation}
\item[$U$-statistic] the Theil's $U$-statistic, presented by
\cite{theil1966}, is a unit free measure obtained as the ratio between the root
MSE (RMSE) of a model and the RMSE of the naive random walk model. A
$U$-statistic less than 1 implies the performance is better than the naive
model.
\end{description}

>>>>

Regularization can be understood using two contexts: learning theory
(probabilistic) and inverse problems (deterministic). In the context of
learning, regularization refers to techniques allowing to avoid over-fitting and
the desire property of the selected estimator is to perform well on new data (to
generalise), e.g. regularized least squares. In the context of inverse problems,
regularization objective is to stabilise, with respect to noise, a possibly
ill-conditioned matrix inversion problem e.g spectral cut-off and Tikhonov
regularization or ridge regression (RR) \cite{ tikhonov1977}. More details in section \ref{app:effcomp}.

@@

\begin{equation}
\label{eq:optsolRR}
\mathbf{\mathbf{X}}_*=(\mathbf{A}^\top \mathbf{A}+\lambda \mathbb{I})^{-1}\mathbf{A}^\top y \, ,
\end{equation}

@@

The major difference between SVM and ANN is the error optimisation. In ANN, the aim of learning is to obtain a set of weight values which minimise the training error while training adjust the capacity of the machine.

@@

In batch learning, the training phase is commonly a computationally expensive
process. Therefore, when new data arrives it can't be included easily into the
model. Moreover, it could happen that we won't have enough time to process the
new data before more data arrives. In online learning algorithms there is no
training phase, the model is updated and evaluated at every time step. This
model updating is computationally less expensive than a training phase.

@@

The objective is to minimise this loss function for all instances.
The quality of online learning algorithms is measured by a quantity known as
regret which is the difference between the performance of the online algorithm
and its optimal predictor $E^* \in \Theta$ given by:

@@

@@

If we define $\displaystyle\mathbf{S}= \sum_{t=1}^{m+1} \mathbf{a}_t
\mathbf{a}_t  ^\intercal + \gamma \mathbb{I} $ and $\mathbf{W}=
\displaystyle\sum_{t=1}^m \mathbf{a}_t \mathbf{y}_t$, the
algorithm~\ref{alg:AAR} is slightly different to the algorithm~\ref{alg:RR}, 
which updated matrix $\mathbf{S}$ before making the prediction.

%Despite the bias-variance tradeoff provides a conceptual framework for
%determining a good model, is not directly useful. Some popular methods
%for determine model selection are:
%
%\begin{itemize}
%    \item Akaike information criterion (AIC) (Akaike, 1974)
%    \item Bootstrap-based selection (Efron and Tibshirani, 1997)
%    \item Cross-validation (Stone, 1974)
%\end{itemize}
%
%On the other hand, $\lambda$ can also be used to reduce computational time as it
%is shown in equation~(\ref{eq:taylor}). Golub (~\cite{golub1965}) suggested that
%$\lambda$ be chosen so that:
%
%\[
%\frac{\lambda}{\lambda + \delta^2} < 0.1
%\]
%
%\noindent where $\delta$ is a lower bound of the smallest non-zero singular
%value $\sigma_k$. One approach is to choose the greatest lower bound for
%$\delta=\sigma_k$. Hence:
%
%\[
%\frac{\lambda}{\lambda + \sigma_k^2} < 0.1
%\]
%
%\noindent if we choose $\lambda=\beta \sigma_k^2$ then we get the following
%relation:
%
%
%\[
%\frac{\beta \sigma_k^2}{\beta \sigma_k^2 + \sigma_k^2} =
%\frac{\beta}{\beta+1} < 0.1 \Rightarrow \beta < \frac{1}{9}  
%\]
%
%Experiments done by ~\cite{coleman+sun2010} have shown that $\beta = 0.01$
%produces satisfactory results. This means that $\lambda = 0.01 \sigma_k^2$.
%
%However, get $\sigma_k$ imply obtaining first the SVD, which is
%computational expensive.
%
%Since the algorithm presented by ~\cite{coleman+sun2010} already compute the QR
%factorization of the matrix $\mathbf{A}$ they show a procedure for getting
%a $\lambda$ approximation using QR.
%
%\begin{enumerate}
%\item Compute QR factorization: $\mathbf{A} = \mathbf{Q_1R_1}$
%\item Let $\mathbf{W}$ denote the set of absolute values of the nonzero diagonal
%elements of $\mathbf{R_1}$. Let $w_{\text{min}}$ and $w_{\text{max}}$ denote the
%smallest and largest elements of $\mathbf{W}$ respectively. Both 
%$\lambda_1 = \hat{\beta} w_{\text{min}}^2$
%$\lambda_2 = \hat{\beta} \frac{w_{\text{min}}^2}{w_{\text{max}}^2}$ where
%$\hat{\beta}=0.00025$ produce satisfactory results.
%\item The $\lambda_{\text{QR}}$ approximation is obtained as:
%$\lambda_{\text{QR}} = \frac{\lambda_1 + \lambda_2}{2}$  
%\end{enumerate}
\newpage

@@

Forecast performance is evaluated using different methods. We have chosen three
measures usually used:
\begin{description}
\item[MAPE] Mean Average Percent Error which presents forecast errors as a
percentage.
\begin{equation}\label{eq:MAPE}
\text{MAPE} = \frac{1}{N} \sum_{t=1}^{N} 
\frac{\left|\mathbf{y}_t-\hat{\mathbf{y}}_t\right|}{\left|\mathbf{y}_t\right|}
 \times 100 
\end{equation}
\item[MAE] Mean Average Error which measures the distance between forecasts to the
true value.
\begin{equation}\label{eq:MAE}
\text{MAE} = \frac{1}{N} \sum_{t=1}^{N} 
\left| 
\mathbf{y}_t-\hat{\mathbf{y}}_t
\right| 
\end{equation}
\item[MSE]  Mean Square Error measures the distance between forecasts
and the true values and large deviations from the true value have a
large impact due to squaring forecast error.
\begin{equation}\label{eq:MSE}
\text{MSE} = 
\frac{\displaystyle \sum_{t=1}^{N} (\mathbf{y}_t-\hat{\mathbf{y}}_t)^2}{N}
\end{equation}
\item[RMSE] Root Mean Square Error also measures the distance between forecasts
to the true values but, unlike MAE, large deviations from the true value have a
large impact on RMSE due to squaring forecast error.
\begin{equation}\label{eq:RMSE}
\text{RMSE} = \sqrt{
\frac{\displaystyle \sum_{t=1}^{N} (\mathbf{y}_t-\hat{\mathbf{y}}_t)^2}{N}}
\end{equation}
\item[$U$-statistic] the Theil's $U$-statistic, presented by
\cite{theil1966}, is a unit free measure obtained as the ratio between the root
MSE (RMSE) of a model and the RMSE of the naive random walk model. 
\end{description}
